Quantile Regressions via Method of Moments with multiple fixed effects

### 2023-10-20 article/article.qmd
s
s
d
d

Quantile regression (QR), introduced by Koenker and Bassett (1978), is an estimation strategy used for modeling the relationships between explanatory variables X and the conditional quantiles of the dependent variable $q_\tau (y|x)$. Using QR one can obtain richer characterizations of the relationships between dependent and independent variables, by accounting for otherwise unobserved heterogeneity.

A relatively recent development in the literature has focused on extending quantile regressions analysis to include individual fixed effects in the framework of panel data. However, as described in Neyman and Scott (1948), and Lancaster (2000), when individual fixed effects are included in quantile regression analysis it generates an incident parameter problem. While many strategies have been proposed for estimating this type of model (see Galvão and Kato, 2018 for a brief review), neither has become standard because of their restrictive assumptions in regards to the individual effects, the computational complexity, and implementation.

More recently, @mss2019 (MSS hereafter) proposed a methodology based on a conditional location-scale model similar to the one described in @he1997, for the estimation of quantile regressions models for panel data via a method of moments. This method allows individual fixed effects allowing to have heterogeneous effects on the entire conditional distribution of the outcome, rather constraining their effect to be a location shift only as in Canay (2011), Koenker(2004), and Lancaster(2000).

In principle, under the assumption that data generating process behind the data is based on a multiplicative heteroskedastic process that is linear in parameters (Cameron and Trivedi, 2005; @mss2019; @he1997), the effect of a variable $X$ on the $q_th$ quantile can be derived as the combination of a location effect, and scale effect moderated by the quantile of an underlying i.i.d. error. For statistical inference, MSS derives the asymptotic distribution of the estimator, suggesting the use of bootstrap standard errors, as well.

@@

In this framework, our paper expands on @mss2019, following some of the suggestions by the authors regarding further research. First, making use of the properties of GMM estimators, we derive various alternatives for the estimation of standard errors based on the empirical Influence functions of the estimators. Second, we reconsider the application of Frisch–Waugh–Lovell (FWL) theorem (Lovell (1963) and Frisch and Waugh’s (1933)) to extend the MSS estimator to allow for the inclusion of multiple fixed effects, for example, individual and year fixed effects.

@@

# Methodology
## Quantile Regression Location-Scale model @he1997

@@

@@

# Appendix {-}

>>>>

Quantile regression (QR), introduced by @koenkerbasset1978, is an estimation strategy used for modeling the relationships between explanatory variables X and the conditional quantiles of the dependent variable $q_\tau (y|x)$. Using QR one can obtain richer characterizations of the relationships between dependent and independent variables, by accounting for otherwise unobserved heterogeneity.

A relatively recent development in the literature has focused on extending quantile regressions analysis to include individual fixed effects in the framework of panel data. However, as described in @neymanscott1948, and @lancaster2000, when individual fixed effects are included in quantile regression analysis it generates an incident parameter problem. While many strategies have been proposed for estimating this type of model (see Galvão and Kato, 2018 for a brief review), neither has become standard because of their restrictive assumptions in regards to the individual effects, the computational complexity, and implementation.

More recently, @mss2019 (MSS hereafter) proposed a methodology based on a conditional location-scale model similar to the one described in @he1997, for the estimation of quantile regressions models for panel data via a method of moments. This method allows individual fixed effects allowing to have heterogeneous effects on the entire conditional distribution of the outcome, rather constraining their effect to be a location shift only as in @canay2011, @koenker2004, and @lancaster2000.

In principle, under the assumption that data generating process behind the data is based on a multiplicative heteroskedastic process that is linear in parameters (@cameron2005; @mss2019; @he1997), the effect of a variable $X$ on the $q_th$ quantile can be derived as the combination of a location effect, and scale effect moderated by the quantile of an underlying i.i.d. error. For statistical inference, MSS derives the asymptotic distribution of the estimator, suggesting the use of bootstrap standard errors, as well.

@@

In this framework, our paper expands on @mss2019, following some of the suggestions by the authors regarding further research. First, making use of the properties of GMM estimators, we derive various alternatives for the estimation of standard errors based on the empirical Influence functions of the estimators. Second, we reconsider the application of Frisch–Waugh–Lovell (FWL) theorem (@frishwaugh1933 and @lovell1963) to extend the MSS estimator to allow for the inclusion of multiple fixed effects, for example, individual and year fixed effects.

@@

# Methodology

## Quantile Regression Location-Scale model @he1997

Quantile regressions are used to identify relationships between the explanatory variables $x$ and the conditional quantiles of the dependent variable $q_\tau (y|x)$. One of the most common formulations to represent the need for quantile regression model is given by:

$$Y=X\beta(\tau), \tau \sim uniform(0,1) $$

Which allows for nonlinearities on the effect of $X$ on $Y$ along all values of $\tau$. This formulation , also known as a random coefficient model, can be re-written as a location-scale model, as described in @he1997:

$$Y = X \beta + X \gamma(\tau)$$

Here, $\beta$ represent the location effect, which represents the average effect $X$ has on $Y$ and can be estimated using e , and $\gamma(\tau)$ represents the scale effect, which can be interpreted as the effect of $X$ on the conditional variance of $Y$.

@@

# Appendix {.appendix}

@@
